\documentclass[a4paper,10pt]{scrartcl}
\RequirePackage[hyperindex]{hyperref}
\usepackage[pdftex]{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\newcommand\cI{{\cal I}}
\newcommand\cT{{\cal T}}
\newcommand\cA{{\cal A}}
\newcommand\cK{{\cal K}}
\newcommand\cC{{\cal C}}
\newcommand\cL{{\cal L}}
\newcommand\fluffy{{\mathit{fluffy}}}

\newenvironment{example}{}{}
\newenvironment{definition}{}{}

\begin{document}
\title{TRILL on SWISH Manual}
\maketitle

\input{syntax}
\input{semantics}
\input{inference}
\input{inf_goals}
\subsection{Executing a Query with TRILL}
To run a query, you can simply load the Prolog file, for example \href{http://trill.lamping.unife.it/example/trill/peoplePets.pl}{\texttt{peoplePets.pl}}, as
\begin{verbatim}
?- [peoplePets].
\end{verbatim}
\noindent
The linked file contains axioms defined in both syntaxes accepted by TRILL, RDF/XML and Prolog Syntax, based on definition of \href{http://vangelisv.github.io/thea/}{Thea} library.
\href{http://trill.lamping.unife.it/example/trill/peoplePets.pl}{\texttt{peoplePets.pl}} is equivalent with the following KB
\begin{verbatim}
:- use_module(library(trill)).

:- trill.

:- add_kb_prefix('','http://cohse.semanticweb.org/ontologies/people#').

subClassOf(someValuesFrom('has_animal', 'pet'), 'natureLover').
subClassOf('cat', 'pet').
annotationAssertion('disponte:probability', 
classAssertion('cat', 'Fluffy'), literal('0.4')).
annotationAssertion('disponte:probability', 
classAssertion('cat', 'Tom'), literal('0.3')).
annotationAssertion('disponte:probability', 
subClassOf('cat', 'pet'), literal('0.6')).
propertyAssertion('has_animal', 'Kevin', 'Fluffy').
propertyAssertion('has_animal', 'Kevin', 'Tom').
classAssertion('cat', 'Fluffy').
classAssertion('cat', 'Tom').
\end{verbatim}
\noindent


You can also load an RDF/XML file using the predicate \verb|load_owl_kb(<filename>).| in the following way:
\begin{itemize}
	\item load TRILL library\\
	\verb|use_module(library(trill)).|
	\item initialize the algorithm you want to perform inference\\
	\verb|init_trill(<algorithm_name>).|\\
	For example, if you want to use TRILL$^P$ you should run \verb|init_trill(trillp).|
	\item load the KB\\
	\verb|load_owl_kb(<filename>).|\\
	For example, \verb|load_owl_kb('./examples/biopaxLevel3_rdf.owl').|
\end{itemize}
\noindent
Now the KB is loaded and the queries can be executed in the usual way.

\section{Download Query Results through an API}
The results of queries can also be downloaded programmatically by directly
approaching the Pengine API. Example client code is 
\href{https://github.com/friguzzi/trill-on-swish/tree/master/client}{available}.  For example, the \verb|swish-ask.sh| client
can be used with bash to download the results for a query in CSV.  The call
below downloads a CSV file for the coin example.
\begin{verbatim}
$ bash swish-ask.sh --server=http://trill.lamping.unife.it \
  examples/trill/peoplePets.pl \
  Prob "prob_instanceOf('natureLover','Kevin',Prob)"
\end{verbatim}
The script can ask queries against Prolog scripts stored in 
\url{http://trill.lamping.unife.it} by specifying
the script on the commandline.  User defined files stored
in TRILL on SWISH (locations of type
\url{http://trill.lamping.unife.it/p/johnEmployee_user.pl}) can
be directly used, for example:
\begin{verbatim}
$ bash swish-ask.sh --server=http://trill.lamping.unife.it \
  johnEmployee_user.pl Expl "instanceOf(person,john,Expl)"
\end{verbatim}
Example programs can be used by specifying the folder portion of the url of the example,
as in the first johnEmployee example above where the url for the program is 
\url{http://trill.lamping.unife.it/examples/trill/johnEmployee.pl}.

You can also use an url for the program as in 
\begin{verbatim}
$ bash swish-ask.sh --server=http://trill.lamping.unife.it \
  https://raw.githubusercontent.com/friguzzi/trill-on-swish/\
  master/examples/trill/peoplePets.pl \
  Prob "prob_instanceOf('natureLover','Kevin',Prob)"
\end{verbatim}
Results can be downloaded in JSON using the option \verb|--json-s| or
\verb|--json-html|.
With the first the output is in a simple string format where Prolog terms are sent using quoted write, the latter serialize responses as HTML strings. E.g.
\begin{verbatim}
$ bash swish-ask.sh --json-s --server=http://trill.lamping.unife.it \
  johnEmployee_user.pl Expl "instanceOf(person,john,Expl)"
\end{verbatim}
The JSON format can also be modified. See
\url{http://www.swi-prolog.org/pldoc/doc_for?object=pengines%3Aevent_to_json/4}.

Prolog can exploit the Pengine API directly.  For example, the above can
be called as:
\begin{verbatim}
?- [library(pengines)].
?- pengine_rpc('http://trill.lamping.unife.it',
     prob_instanceOf('natureLover','Kevin',Prob),
     [ src_url('https://raw.githubusercontent.com/friguzzi/trill-on-swish/\
  master/examples/trill/peoplePets.pl'),
     application(swish)
     ]).
Prob = 0.51.
?-
\end{verbatim}

\section{Manual in PDF}
A PDF version of the manual is available at
\url{https://github.com/rzese/trill/blob/master/doc/help-trill.pdf}.


\bibliographystyle{plain}
\bibliography{bib}

\appendix
\input{appendix}
\end{document}