\section{Syntax}
\label{syn}

Description Logics (DLs) are knowledge representation formalisms that are at the basis of the Semantic Web \cite{DBLP:conf/dlog/2003handbook,dlchap} and are used for modeling ontologies.
They are represented using a syntax based on concepts, basically sets of individuals of the domain, and roles, sets of pairs of individuals
of the domain. In this section, we recall the expressive description logic $\mathcal{ALC}$ \cite{DBLP:journals/ai/Schmidt-SchaussS91}. We refer to 
\cite{DBLP:journals/ws/LukasiewiczS08} for a detailed description of $\mathcal{SHOIN}(\mathbf{D})$ DL, that is at the basis of OWL DL.

%%ALC
Let $\mathbf{A}$, $\mathbf{R}$ and $\mathbf{I}$ be sets of \emph{atomic concepts}, \emph{roles} and \emph{individuals}.
A \emph{role} is an atomic role $R \in \mathbf{R}$. 
\emph{Concepts} are defined by induction as follows. Each $C \in \mathbf{A}$, $\bot$ and $\top$
are concepts.
If $C$, $C_1$ and $C_2$ are concepts and $R \in \mathbf{R}$, then $(C_1\sqcap C_2)$, $(C_1\sqcup C_2 )$, $\neg C$,
$\exists R.C$, and $\forall R.C$ are concepts. 
Let $C$, $D$ be concepts,  $R \in \mathbf{R}$ and $a, b \in \mathbf{I}$. 
An \emph{ABox} $\cA$ is a finite set of \textit{concept membership axioms} $a : C$ and \textit{role membership
axioms} $(a, b) : R$, while 
a \emph{TBox} $\cT$ is a finite set of \textit{concept inclusion axioms} $C\sqsubseteq D$. $C \equiv D$ abbreviates $C \sqsubseteq D$ and $D\sqsubseteq  C$.

A \emph{knowledge base} $\cK = (\cT, \cA)$ consists of a TBox $\cT$ and an ABox $\cA$.
A KB $\cK$ is assigned a semantics in terms of set-theoretic interpretations $\cI = (\Delta^\cI , \cdot^\cI )$, where $\Delta^\cI$ is a non-empty \textit{domain} and $\cdot^\cI$ is the \textit{interpretation function} that assigns  an element in $\Delta ^\cI$ to each $a \in \mathbf{I}$, a subset of $\Delta^\cI$ to each $C \in \mathbf{A}$ and a subset of $\Delta^\cI \times \Delta^\cI$ to each $R \in \mathbf{R}$.

TRILL allows the use of two different syntaxes used together or individually:
\begin{itemize}
 \item RDF/XML
 \item TRILL syntax 
\end{itemize}

RDF/XML syntax can be used by exploiting the predicate \verb|owl_rdf/1|. For example:
\begin{verbatim}
owl_rdf('
<?xml version="1.0"?>

<!DOCTYPE rdf:RDF [
    <!ENTITY owl "http://www.w3.org/2002/07/owl#" >
    <!ENTITY xsd "http://www.w3.org/2001/XMLSchema#" >
    <!ENTITY rdfs "http://www.w3.org/2000/01/rdf-schema#" >
    <!ENTITY rdf "http://www.w3.org/1999/02/22-rdf-syntax-ns#" >
]>

<rdf:RDF xmlns="http://here.the.IRI.of.your.ontology#"
     xml:base="http://here.the.IRI.of.your.ontology"
     xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
     xmlns:owl="http://www.w3.org/2002/07/owl#"
     xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
     xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#">
    <owl:Ontology rdf:about="http://here.the.IRI.of.your.ontology"/>

    <!-- 
    Axioms
    -->

</rdf:RDF>
').
\end{verbatim}
For a brief introduction on RDF/XML syntax see \href{rdfxml-syn}{\texttt{RDF/XML syntax and tools}} section below.

Note that each single \verb|owl_rdf/1| must be self contained and well formatted, it must start and end with \verb|rdf:RDF| tag and contain all necessary declarations (namespaces, entities, ...).


An example of the combination of both syntaxes is shown the example \href{http://trill.lamping.unife.it/example/trill/johnEmployee.pl}{\texttt{johnEmployee.pl}}. It models that \textit{john} is an \textit{employee} and that employees are \textit{workers}, which are in turn people (modeled by the concept \textit{person}).
\begin{verbatim}
owl_rdf('<?xml version="1.0"?>
<rdf:RDF xmlns="http://example.foo#"
     xml:base="http://example.foo"
     xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
     xmlns:owl="http://www.w3.org/2002/07/owl#"
     xmlns:xml="http://www.w3.org/XML/1998/namespace"
     xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
     xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#">
    <owl:Ontology rdf:about="http://example.foo"/>

    <!-- Classes -->
    <owl:Class rdf:about="http://example.foo#worker">
        <rdfs:subClassOf rdf:resource="http://example.foo#person"/>
    </owl:Class>

</rdf:RDF>').

subClassOf('employee','worker').

owl_rdf('<?xml version="1.0"?>
<rdf:RDF xmlns="http://example.foo#"
     xml:base="http://example.foo"
     xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
     xmlns:owl="http://www.w3.org/2002/07/owl#"
     xmlns:xml="http://www.w3.org/XML/1998/namespace"
     xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
     xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#">
    <owl:Ontology rdf:about="http://example.foo"/>
    
    <!-- Individuals -->
    <owl:NamedIndividual rdf:about="http://example.foo#john">
        <rdf:type rdf:resource="http://example.foo#employee"/>
    </owl:NamedIndividual>
</rdf:RDF>').
\end{verbatim}

\subsection{TRILL Syntax}
\label{trill-syn}
\subsubsection{Declarations}


TRILL syntax allows, as in standard OWL, the declaration of classes, properties, etc.
\begin{verbatim}
class("classIRI").
datatype("datatypeIRI").
objectProperty("objectPropertyIRI").
dataProperty("dataPropertyIRI").
annotationProperty("annotationPropertyIRI").
namedIndividual("individualIRI").
\end{verbatim}
However, TRILL properly works also in their absence.

TRILL syntax allows also the declaration of aliases for namespaces by using the \verb|kb_prefix/2| predicate.
\begin{verbatim}
kb_prefix("foo","http://example.foo#").
\end{verbatim}
After this declaration, the prefix \verb|foo| is available, thus, instead of \verb|http://example.foo#john|, one can write \verb|foo:john|.
It is possible to define also an empty prefix as
\begin{verbatim}
kb_prefix("","http://example.foo#").
\end{verbatim}
or as
\begin{verbatim}
kb_prefix([],"http://example.foo#").
\end{verbatim}
In this way \verb|http://example.foo#john| can be written only as \verb|john|.

\textbf{Note:} Only one prefix per alias is allowed. Aliases defined in OWL/RDF part have the precedence, in case more than one prefix was assigned to the same alias, TRILL keeps only the first assignment.


\subsubsection{Axioms}


Axioms are modeled using the following predicates
\begin{verbatim}
subClassOf("subClass","superClass").
equivalentClasses([list,of,classes]).
disjointClasses([list,of,classes]).
disjointUnion([list,of,classes]).

subPropertyOf("subProperty","superProperty").
equivalentProperties([list,of,properties]).
propertyDomain("propertyIRI","domainIRI").
propertyRange("propertyIRI","rangeIRI").
transitiveProperty("propertyIRI").

sameIndividual([list,of,individuals]).
differentIndividuals([list,of,individuals]).

classAssertion("classIRI","individualIRI").
propertyAssertion("propertyIRI","subjectIRI","objectIRI").
annotationAssertion("annotationIRI",axiom,literal('value')).
\end{verbatim}
For example, for asserting that \textit{employee} is subclass of \textit{worker} one can use
\begin{verbatim}
subClassOf(employee,worker).
\end{verbatim}
while the assertion \textit{worker} is equal to \textit{workingman} can be defined as
\begin{verbatim}
equivalentClasses([worker,workingman]).
\end{verbatim}


Annotation assertions can be defined, for example, as
\begin{verbatim}
annotationAssertion(foo:myAnnotation,subClassOf(employee,worker),'myValue').
\end{verbatim}


In particular, an axiom can be annotated with a probability which defines the degree of belief in the truth of the axiom. See \cite{trillquale} for details.


Below, an example of an probabilistic axiom, following the TRILL syntax.
\begin{verbatim}
annotationAssertion('disponte:probability',subClassOf(employee,worker),literal('0.6')).
\end{verbatim}

\subsubsection{Concepts descriptions}
Complex concepts can be defined using different operators.

Existential and universal quantifiers
\begin{verbatim}
someValuesFrom("propertyIRI","classIRI").
allValuesFrom("propertyIRI","classIRI").
\end{verbatim}
Union and intersection of concepts
\begin{verbatim}
unionOf([list,of,classes]).
intersectionOf([list,of,classes]).
\end{verbatim}
Cardinality descriptions
\begin{verbatim}
exactCardinality(cardinality,"propertyIRI").
exactCardinality(cardinality,"propertyIRI","classIRI").
maxCardinality(cardinality,"propertyIRI").
maxCardinality(cardinality,"propertyIRI","classIRI").
minCardinality(cardinality,"propertyIRI").
minCardinality(cardinality,"propertyIRI","classIRI").
\end{verbatim}
Complement of a concept
\begin{verbatim}
complementOf("classIRI").
\end{verbatim}
Nominal concept
\begin{verbatim}
oneOf([list,of,classes]).
\end{verbatim}
For example, the class \textit{workingman} is the intersection of \textit{worker} with the union of \textit{man} and \textit{woman}. It can be defined as:
\begin{verbatim}
equivalentClasses([workingman,intersectionOf([worker,unionOf([man,woman])])]).
\end{verbatim}

\subsection{RDF/XML syntax and tools}
\label{rdfxml-syn}
As said before, TRILL is able to automatically translate RDF/XML knowledge bases when passed as a string using 
the preticate \verb|owl_rdf/1|.

Consider the following axioms 

\begin{verbatim}
classAssertion(Cat,fluffy)
subClassOf(Cat,Pet)
propertyAssertion(hasAnimal,kevin,fluffy)
\end{verbatim}

The first axiom states that \textit{fluffy} is a \textit{Cat}. The second states that every \textit{Cat} is also a \textit{Pet}. The third states that the role \textit{hasAnimal} links together \textit{kevin} and \textit{fluffy}.

RDF (Resource Descritpion Framework) is a standard W3C. See the \href{http://www.w3.org/TR/REC-rdf-syntax/}{syntax specification} for more details.
RDF is a standard XML-based used for representing knowledge by means of triples.
A representations of the three axioms seen above is shown below.
\begin{verbatim}
<owl:NamedIndividual rdf:about="fluffy">
  <rdf:type rdf:resource="Cat"/>
</owl:NamedIndividual>

<owl:Class rdf:about="Cat">
  <rdfs:subClassOf rdf:resource="Pet"/>
</owl:Class>

<owl:ObjectProperty rdf:about="hasAnimal"/>
<owl:NamedIndividual rdf:about="kevin">
 <hasAnimal rdf:resource="fluffy"/>
</owl:NamedIndividual>
\end{verbatim}

Annotations are assertable using an extension of RDF/XML. For example the annotated axiom below, defined using the TRILL sintax
\begin{verbatim}
annotationAssertion('disponte:probability',subClassOf('Cat','Pet'),literal('0.6')).
\end{verbatim}
is modeled using RDF/XML syntax as
\begin{verbatim}
<owl:Class rdf:about="Cat">
 <rdfs:subClassOf rdf:resource="Pet"/>
</owl:Class>
<owl:Axiom>
 <disponte:probability rdf:datatype="&amp;xsd;decimal">0.6</disponte:probability>
 <owl:annotatedSource rdf:resource="Cat"/>
 <owl:annotatedTarget rdf:resource="Pet"/>
 <owl:annotatedProperty rdf:resource="&amp;rdfs;subClassOf"/>
</owl:Axiom>
\end{verbatim}
If you define the annotated axiom in the RDF/XML part, the annotation must be declared in the knowledge base as follow
\begin{verbatim}
<!DOCTYPE rdf:RDF [
 ...
 <!ENTITY disponte "https://sites.google.com/a/unife.it/ml/disponte#" >
]>

<rdf:RDF
 ...
 xmlns:disponte="https://sites.google.com/a/unife.it/ml/disponte#"
 ...>

 ...
 <owl:AnnotationProperty rdf:about="&amp;disponte;probability"/>
 ...
</rdf:RDF>
\end{verbatim}

There are many \href{http://www.w3.org/2001/sw/wiki/Category:Editor}{editors} for developing knowledge bases.
