\documentstyle[11pt,prospectra,openbib]{article}
%\documentstyle[11pt,prospectra,openbib,adbgerman]{article}

\include{layout}
\include{macros}

\begin{document}

\title{
{\bf CEC}\\
A System to Support Modular Order-Sorted Specifications
with Conditional Equations\\
User Manual (Version 1.5)}
\author{
Hubert Bertling\\
Harald Ganzinger\\
Renate Sch\"afers}

\prospectrano{M.1.3--R--18.0}
\prospectradist{public}

\maketitle

\begin{abstract}
{\bf CEC} is a rewrite rule laboratory for order-sorted specifications
with conditional equations. The major module of CEC is a powerful
completion procedure for conditional equations. This manual describes its 
use assuming the user to be familiar with basic notions in
conditional term rewriting.
\end{abstract}

\setcounter{page}{1}
\pagenumbering{roman}
\tableofcontents

\newpage

\setcounter{page}{1}
\pagenumbering{arabic}


\input{intro}   % --.4.89
\input{example}
\input{syntax}   % --.4.89
\input{inout}   % 2.5.89
\input{display}  % 16.3.89
\input{specops}   % 2.5.89
\input{completion}  % 16.3.89
\input{computing}   % 2.5.89

\input{commands} % 15.3.89


\newpage

\bibliography{lit}
\bibliographystyle{mhshort}
\newpage

\appendix
\section{PREDEFINED OPERATORS OF CEC}
\label{PredefinedOperators}
\begin{center}
\begin{tabular}{|lrrcc|} \hline
CEC operators & priority & fix & reserved & changable \\ \hline
\cec{=>} & 950 & xfx & + & - \\
\cec{=} & 700 & xfx & + & - \\
\cec{:} & 600 & xfy & + & - \\
\cec{and} & 850 & fy & + & - \\
\cec{@} & 50 & fx & + & - \\ \hline
\cec{->} & 1050 & xfy & - & - \\ 
\cec{cons}, \cec{op} & 950 & fx & - & - \\
\cec{in} & 935 & xfy & - & - \\
\cec{let} & 910 & fy & - & - \\
\cec{var} & 1100 & fx & - & - \\
\cec{<} & 700 & xfx & - & - \\
\cec{using} & 600 & xfx & - & - \\
\cec{module}, \cec{order} & 500 & fx & - & - \\
\cec{+} & 500 & yfx & - & - \\
\cec{for} & 400 & xfx & - & - \\
\cec{*} & 400 & yfx & - & - \\ \hline
\cec{:-}, \cec{-->} & 1200 & xfx & - & + \\
\cec{:-}, \cec{?-}  & 1200 & fx  & - & + \\
\cec{public}, \cec{multifile}, \cec{mode}, \cec{meta_predicate}, \cec{dynamic} & 1150 & fx & - & + \\
\cec{;}             & 1100 & xfy & - & + \\
\cec{,}             & 1000 & xfy & - & + \\
\cec{spy}, \cec{nospy}, \cec{\+} & 900 & fy & - & + \\
\cec{not} & 900 & fx & - & + \\
\cec{is}, \cec{=..}, \cec{==}, \cec{\==}, 
\cec{@<}, \cec{@>}, \cec{@=<}, \cec{@>=}, \cec{=:=}, \cec{=\=},
\cec{>}, \cec{=<}, \cec{>=} & 700 & xfx & - & + \\
\cec{-}, \cec{/\}, \cec{\/} & 500 & yfx & - & + \\
\cec{|} & 500 & xfy & - & + \\
\cec{-} & 500 & fx & - & + \\
\cec{/}, \cec{//}, \cec{<<}, \cec{>>} & 400 & yfx & - & + \\
\cec{mod} & 300 & xfx & - & + \\
\cec{^} & 200 & xfy & - & + \\ 
\cec{?} & 200 & fx & - & + \\
\hline
\end{tabular}
\end{center}

If an operator is {\em reserved}, it cannot me used in any signature. If 
an operator is not {\em changable}, it can be used in a signature, but 
its priority and fix may not be changed. 

\section{INSTALLATION OF CEC}
The CEC distribution tape contains one tar file, which
includes the CEC-System. The size of the CEC sources is about 763 kbytes, the whole system needs
about  7657 kbytes.
To install CEC execute the following steps:\bigskip

\noindent
Create a directory ``\cec{cec}'' where you want to file in CEC, 
and use the tar-command to read the tar file.

\noindent
Then change to the directory ``\cec{cec/obj/cec}'' and correct
the value of environment variables\bigskip

\cec{PROSPECTRA}

\cec{QPROLOG}\bigskip

\noindent
in the file ``\cec{Makefile}'' before calling\bigskip

{\bf make cec}\bigskip

\noindent
After successful creation of a executable \cec{cec} call\bigskip

{\bf make clean}

{\bf make install}\bigskip

\noindent
to complete the installation of CEC.\bigskip

\noindent
The executable file is located in the directory ``\cec{cec/bin}'' and
you can work with CEC as specified in the user manual, e.g. \bigskip

\verb+| ?-+ {\bf cd 'demo/cec/math'.}

\verb+| ?-+ {\bf in(abelianGroup,poly1).}

\verb+| ?-+ {\bf c.}

\end{document}
